 

### 2021-01-27 vertical-thermal-review.Rmd
bold font on topic sentences (intro)

Globally, temperatures have risen # $^\circ$C since YEAR and are expected to rise an additional #-#$^\circ$C by YEAR (IPCC REF), accompanied by increasing severity and frequency of heat waves [@perkinsIncreasingFrequencyIntensity2012] and hotter droughts [REFS].
These changes are expected to have profound effects on tree metabolism and forest ecosystem function (REFS; IPBES report), including *(altered rates of photosynthesis and respiration)* [REF], foliar damage during heat waves [], and reduced growth and elevated mortality during drought [REFS].
Individually or in combination, these changes have the potential to reduce forest C sequestration, as may already be occurring in some forests [REFS; @sullivanLongtermThermalSensitivity2020]. 
The resulting feedbacks to the climate system (e.g., decreased C storage, altered albedo and hydrology) will in turn have a critical influence on the future trajectory of climate change [@bonan_ecological_2016, Bonan 2008], yet this influence remains extremely uncertain [REFS].

@@

**We lack a systematic, up to date understanding of biophysical and biological patterns across the vertical canopy gradient in forests, how this gradient affects leaf-level processes, and in turn how it affects ecosystems (Fig. `r fig_schematic`).**

**This review addresses the following questions:**

>>>>

**Forest responses and feedbacks to climate change will have a critical influence on the future of Earth's climate.**
Globally, temperatures have risen # $^\circ$C since YEAR and are expected to rise an additional #-#$^\circ$C by YEAR (IPCC REF), accompanied by increasing severity and frequency of heat waves [@perkinsIncreasingFrequencyIntensity2012] and hotter droughts [REFS].
These changes are expected to have profound effects on tree metabolism and forest ecosystem function (REFS; IPBES report), including *(altered rates of photosynthesis and respiration)* [REF], foliar damage during heat waves [], and reduced growth and elevated mortality during drought [REFS].
Individually or in combination, these changes have the potential to reduce forest C sequestration, as may already be occurring in some forests [REFS; @sullivanLongtermThermalSensitivity2020]. 
The resulting feedbacks to the climate system (e.g., decreased C storage, altered albedo and hydrology) will in turn have a critical influence on the future trajectory of climate change [@bonan_ecological_2016, Bonan 2008], yet this influence remains extremely uncertain [REFS].

@@

**Here we review biophysical and biological patterns across the vertical canopy gradient in forests, how these gradients affect leaf-level processes and, in turn, whole plant performance (Fig. `r fig_schematic`), and the implications for understanding forest ecosystem responses and feedbacks to climate change .**
